Excuse me, do you have a moment to talk about version control?

### 2017-05-30 README.Rmd
Act on many items from Hadley's review + more

Why would a statistician use version control, such as Git? And what is the point
of hosting your work online, e.g., on GitHub? Could the gains possibly justify 
the inevitable pain?

@@

  * Doing your work becomes tightly integrated with organizing, recording, and disseminating it. It's not a separate, burdensome task you are tempted to neglect.
  * Collaboration is much more structured, with powerful tools for asynchronous work and managing versions.
  * The marginal effort required to create a web presence for a project is negligible.
  * GitHub makes a fantastic course management system for courses that use R. You and your students can exchange actual working code and explore the associated results.
  * By using common mechanics across work modes (research, teaching, analysis), you achieve basic competence quickly and avoid the demoralizing forget-relearn cycle.

@@

In my opinion, __for new users__, the pros of Git only outweigh the cons when 
you factor in the overhead of communicating and collaborating with other people,
including your future self. And who among us does not need to do that? Your life
is much easier if this is baked into your workflow, as opposed to being a 
separate process that you dread and avoid. Communication and collaboration are 
the killer apps of version control. Git enforces a rigorous model of file 
management, but it is critical for the distribution of files across different 
people, computers, and time.

This has an implication for selecting your first Git projects: in order to see 
substantial gain for your pain, you need to pick projects that require you to 
share rapidly evolving files with others. It is tempting to pick a quiet little 
project that you work on privately, but you risk missing the main point of 
formal version control.

@@

The Git way is to track the evolution of that file, through a series of commits,
each equipped with an explanatory message and a nickname. All collaborators sync
regularly to a common version, acknowledging that the difficulty of merging goes
up faster than the size of the difference [@good-enough]. Especially important 
versions get a human-readable tag, to signal a meaningful milestone. Yes, there 
is some pain in adopting the formalism of Git, but it is worth it.

@@

This article does not provide step-by-step instructions on how to use Git and 
GitHub. This format would not be effective, but we do provide annotated links to
such resources in ?the appendix or online supplement?. Instead, I'll convey what
the workflow feels like and what the payoffs are, with special attention to the 
statistics and R context. The goal is to help the Git-curious generate the 
activation energy needed to get started.

@@

[GitHub](https://github.com) is currently the most popular Git hosting service. 
Others include [GitLab](https://about.gitlab.com) and 
[Bitbucket](https://bitbucket.org). These services provide a home for Git-based 
projects on the internet. It is a bit like DropBox or Google Drive, but more 
structured, powerful, and programmatic.

The remote host acts as the clearinghouse for a Git-managed project. This allows
other people to see the project files, sync up with the current version, and 
perhaps even make or propose changes. These hosting providers offer 
well-designed web interfaces that are a dramatic improvement over traditional 
Unix Git servers. Many operations can be done entirely in the browser, such as 
viewing current or past versions of a file, commenting on a recent change, and 
editing or adding files. These hosts also offer granular control over who can 
see, edit, and administer a project.

@@

Here is the initial setup process, which you do once or, for some steps, once per computer.

@@

* Install Git. Depending on your OS, Git might already be installed. But many of
us may need to install it or will choose to update to a more recent version.
Some basic configuration is critical, such as setting your username and email.

@@

* Confirm, with a practice repository, that local Git can send and receive the 
current version of the repository on GitHub, known as __pushing__ and __pulling__, respectively. This will require authenticating yourself with GitHub from a local shell or Git client. At this point, most people elect to do a bit of extra setup to ensure that they are not repeatedly challenged for their GitHub credentials going forward.

Once this setup is done, you are ready to start using Git and GitHub with your 
projects.

## Is this going to hurt?

Yes.

Git was built neither for the exact usage described here, nor for broad 
usability. You will undoubtedly notice this, so it can be helpful to know this 
in advance. Happily there are many helpful tools that mediate your interactions 
with Git. GitHub itself is a fine example, as is RStudio. In addition to pointing out tools that soften Git's sharpest edges, I recommend specific habits and attitudes that reduce frustration.

General recommendations for agony reduction:

@@

Recall that a repository or repo is just a directory of files that Git will manage holistically. A commit functions like a snapshot of all the files in the repo, at a specific moment. Under the hood, that is not exactly how Git implements things. Mental models don't have to be accurate in order to be useful, but in this case there's some value in aligning the two.

@@

**Will it play especially well with Git/GitHub?** This boils down to whether Git diffs will be informative and whether GitHub has nice handling for the file type. Small-to-medium plain text files with hard line breaks are a perfect fit, but there are a few more pleasant surprises.

@@

**Will it cause problems with Git/GitHub?** This boils down to whether Git diffs are informative, whether GitHub can render the file usefully, and the file's likely effect on Git operations. A file that is large and changing quite often can make your repository bloated and slow down pushes and pulls. If a file is binary, such as a Word document or Excel spreadsheet, you will not get human-readable diffs, nor can GitHub display it in the browser. Html is not rendered on GitHub, but is, instead, displayed in raw form. Binary files are also a prime source of merge conflicts, because they are beyond the reach of Git's sophisticated automatic merging logic. For many of us, adoption of Git/GitHub should prompt a pivot away from `.docx` and `.pdf` as primary output formats and towards `.md`, at least during periods of rapid development.

@@

Collaboration is probably the most compelling reason to manage a project with Git and GitHub. I have a broad definition of collaboration. It includes hands-on participation by multiple people as well as an asymmetric model, in which some people are active makers and others only read or review.

@@

*something I would like to write about, but have no space for here: different models of collaboration, i.e. everyone working on one repo & one branch vs. making pull requests from branches in the main repo or from forks*

@@

>>>>

Why would a statistician use version control, such as [Git](http://git-scm.com)? And what is the point of hosting your work online, e.g., on [GitHub](https://github.com)? Could the gains possibly justify the inevitable pain?

@@

  * Doing your work becomes tightly integrated with organizing, recording, and disseminating it. It's not a separate, burdensome task you are tempted to neglect.
  * Collaboration is much more structured, with powerful tools for asynchronous work and managing versions.
  * The marginal effort required to create a web presence for a project is negligible.
  * GitHub makes a fantastic course management system for courses that use R. You can exchange actual working code with your students and explore the associated results.
  * By using common mechanics across work modes (research, teaching, analysis), you achieve basic competence quickly and avoid the demoralizing forget-relearn cycle.

Now the bad news: Git was built neither for the exact usage described here, nor for broad usability. You will undoubtedly notice this, so it's best to know in advance. Happily, there are many helpful tools that mediate your interactions with Git. GitHub itself is a fine example, as is [RStudio](https://www.rstudio.com/products/rstudio/). In addition to pointing out tools that soften Git's sharpest edges, I recommend specific habits and attitudes that reduce frustration.

@@

In my opinion, __for new users__, the pros of Git only outweigh the cons when 
you consider the overhead of working with other people, including your future self. And who among us does not need to do that? In a Git-based workflow, you document and, optionally, expose your work as you go. Communication and collaboration are the killer apps of version control. Git's model of file management can feel uncomfortably rigid, but it enables the distribution of files across different people, computers, and time.

This has an implication for selecting your first Git projects: you will enjoy the most gain for your pain if you pick a project that involves sharing rapidly evolving files with others. It is tempting to pick a quiet, private project, but you risk missing out on the main benefits of formal version control.

@@

The Git way is to track the evolution of that file, through a series of commits,
each equipped with an explanatory message and a nickname. All collaborators sync
regularly to a common version, acknowledging that the difficulty of merging only goes up as more changes accumulate [@good-enough]. Especially important 
versions get a human-readable tag, to signal a meaningful milestone. Yes, there 
is some pain in adopting the formalism of Git, but it is worth it.

@@

This article does not provide step-by-step instructions on how to use Git and 
GitHub. This format would not be effective, but annotated links to
such resources are given in the supplementary materials. Instead, I'll convey what the workflow feels like and what the payoffs are, with special attention to the statistics and R context. The goal is to help the Git-curious generate the 
activation energy needed to get started.

@@

Git offers a powerful structure for file management and [GitHub](https://github.com) complements it with a slick user interface and distribution mechanism for Git repositories. GitHub is currently the most popular Git hosting service, providing a home for Git-based projects on the internet. It is a bit like DropBox or Google Drive, but more structured, powerful, and programmatic.

The remote host acts as the clearinghouse for a Git-managed project. This allows
others to see the project files, sync up with the current version, and perhaps even propose or make changes. GitHub has a well-designed web interface that is a dramatic improvement over traditional Unix Git servers. Many operations can be done entirely in the browser, such as viewing current or past versions of a file, commenting on a recent change, and editing or adding files. GitHub also offers granular control over who can see, edit, and administer a project.

@@

Here is an overview of the initial setup process, which happens once or, for some steps, once per computer. This is excerpted from [Happy Git and GitHub for the useR](http://happygitwithr.com), which holds battle-tested instructions honed over several years in [STAT 545](http://stat545.com) at the University of British Columbia.

@@

* Install Git. Depending on your OS, Git might already be installed. But many may need to install it or will choose to update to a more recent version.
Some basic configuration is critical, such as setting your username and email.

@@

* Confirm, with a practice repository, that local Git can send and receive the 
current version of the repository on GitHub, known as __pushing__ and __pulling__, respectively.

Once this setup is done, you are ready to start using Git and GitHub with your 
projects. Some general recommendations for agony reduction:

@@

Recall that a repository or repo is just a directory of files that Git manages holistically. A commit functions like a snapshot of all the files in the repo, at a specific moment. Under the hood, that is not exactly how Git implements things. Mental models don't have to be accurate in order to be useful, but in this case there's some value in aligning the two.

@@

**Will it play nicely with Git/GitHub?** This boils down to whether Git diffs will be informative and whether GitHub has nice handling for the file type. Small-to-medium plain text files with hard line breaks are ideal, but there are a few more pleasant surprises.

@@

**Will it cause problems with Git/GitHub?** This boils down to whether Git diffs are informative, whether GitHub can render the file usefully, and the file's likely effect on Git operations. A file that is large and changing often can make your repository bloated and slow down pushes and pulls. If a file is binary, such as a Word document or Excel spreadsheet, you will not get human-readable diffs, nor can GitHub display it in the browser. Html is not rendered on GitHub, but is, instead, displayed in raw form. Binary files are also a prime source of merge conflicts (see below), because they are beyond the reach of Git's sophisticated automatic merging logic. This all implies that adoption of Git/GitHub should prompt a pivot away from `.docx` and `.pdf` as primary output formats and towards `.md`, at least during periods of rapid development.

@@

Collaboration is the most compelling reason to manage a project with Git and GitHub. My definition of collaboration includes hands-on participation by multiple people as well as an asymmetric model, in which some people are active makers and others only read or review.

@@

@@

instructions re: supplementary materials
http://amstat.tfjournals.com/supplementary-materials/

TAS author instructions
http://www.tandfonline.com/action/authorSubmission?journalCode=utas20&page=instructions
